MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
full paper pass

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The materials of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing  parts made of conventional materials such as aluminum or steel. CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. 
The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or are barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.

Figure~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}.

% Fiber bundles determine to a large extent the properties in terms of specific stiffness, specific strength, or durability of the final component. However, when scanning reinforced polymers using modern XCT devices, there is still a trade-off between view port and image resolution. Due to the intrinsic concept of cone beam and fan beam XCT setups, the magnification reached within a scan is determined by the used distances between source and specimen as well as source and detector. Furthermore the magnification directly influences both resolution and viewport: Higher resolutions decrease the viewport or how much of the structure can be imaged while lower resolutions allow for larger view ports.
%\begin{figure}
%\begin{center}
%\includegraphics[width=0.3\textwidth]{imagesMetaTracts/crop-13-total}
%\caption{Volume rendering of a part of a dataset. Fig~\ref{fig:slice_data:a} and ~\ref{fig:slice_data:b} show image slices along the blue and red plane respectively.  }\label{fig:orig_data}
%\end{center}
%\end{figure}
%!!!!! TODO REFINE DTI PARAGRAPH !!!!!
%Diffusion Tensor Imaging tractography clustering and visualization.
%The topic of DTI visualization has seen an extensive amount of investigation by researchers.
%Advantage of two phase clustering 
%What is the  motivation for fiber-lets
%Seed point selection 
%distance between tracks 
%how well do we do in noise , may be data set 1

%!!!!! I have replaced this with the following para !!!!!
%In this work, we present a framework which interprets and advances these techniques in order to extract and visualize fiber bundles in industrial 3D X-Ray computed tomography data of woven fiber reinforced composites. The main goal of this work is to expand the state of the art in nondestructive testing of composite structures towards lower resolutions and thus larger areas of woven fabrics.

@@

While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}[noitemsep]
\item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
\item{ Which fiber bundles show a particular orientation? }
\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
\end{itemize}

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[h]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.5\textwidth]{images_pvis/workflow.png}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Figure~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see section~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see section~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see section ~\ref{sec:vis}. Section~\ref{sec:results} describes our experimental results, Section~\ref{sec:param_choices} presents our parameter choices and  Section~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.

@@

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

There are a number of difficulties in hierarchical clustering. First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores. Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.

Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=0.5\textwidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.

We make the following assumptions on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\section {Extracting MetaTracts }
\label{sec:approach}
Extracting the MetaTracts consists of two  major steps. First we extract a local orientation vector at each grid location (section~\ref{subsec:rh}). Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (section~\ref{subsec:mtprop}).

@@

Grid locations where the $R_{H}$ is above a certain threshold are marked as regions with reliable orientations.
We use 0.3 as the cutoff for all our test cases. (See Section~\ref{sec:param_choices} on the choice of 0.3).

@@

In Figure~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. Figure~\ref{fig:reliable_hessian}a shows the entire data set. Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. Figure~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively. Figure~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=0.5\textwidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
\label{fig:reliable_hessian}
\end{figure}

@@

 %glass
We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
Dataset 1 is described in section~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which we identified correctly.
Figure~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).

>>>>

\section{Introduction and Motivation}
\label{sec:intro}
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and components. The material of choice for a growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing parts made of conventional materials such as aluminum or steel. But the application areas of CFRPs are not limited on these two industries. For a wide range of different areas ranging from marine and civil engineering, wind energy, to sporting goods manufacturing, CFRP is highly interesting. The main reason therefore is found in CFRP materials having desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. CFRP materials show these characteristics at considerably lower weight compared to conventional materials. At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs. Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates we are addressing in this work, consist of two main components:
a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the properties of the CFRP laminate. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. After curing, the production process of the CFRP laminate is finished.

The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work, we focus on datasets with larger viewports and lower resolutions where the individual carbon fibers (filaments) are indiscernible or barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of individual fibers.
%of short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Fig.~\ref{fig:data-char} provides an example of our targeted dataset. It also shows the recurring fiber bundle pattern in the final CFRP laminate, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On one hand, CFRP components have gained wide application in industry because of its superior material and physical properties in comparison to conventional materials~\cite{Karpat2012}. On the other hand, recent developments of industrial 3D X-ray computed tomography (XCT) with regard to larger detectors, larger field of views, and better resolutions opened XCT for this new application area of non destructive testing for fiber reinforced components~\cite{Schilling2005}. 
% Fiber bundles determine to a large extent the properties in terms of specific stiffness, specific strength, or durability of the final component. However, when scanning reinforced polymers using modern XCT devices, there is still a trade-off between view port and image resolution. Due to the intrinsic concept of cone beam and fan beam XCT setups, the magnification reached within a scan is determined by the used distances between source and specimen as well as source and detector. Furthermore the magnification directly influences both resolution and viewport: Higher resolutions decrease the viewport or how much of the structure can be imaged while lower resolutions allow for larger view ports.
%\begin{center}
%\includegraphics[width=0.3\textwidth]{imagesMetaTracts/crop-13-total}
%\caption{Volume rendering of a part of a dataset. Fig~\ref{fig:slice_data:a} and ~\ref{fig:slice_data:b} show image slices along the blue and red plane respectively.  }\label{fig:orig_data}
%\end{center}
%!!!!! TODO REFINE DTI PARAGRAPH !!!!!
%Diffusion Tensor Imaging tractography clustering and visualization.
%The topic of DTI visualization has seen an extensive amount of investigation by researchers.
%Advantage of two phase clustering 
%What is the  motivation for fiber-lets
%Seed point selection 
%distance between tracks 
%how well do we do in noise , may be data set 1
%!!!!! I have replaced this with the following para !!!!!
%In this work, we present a framework which interprets and advances these techniques in order to extract and visualize fiber bundles in industrial 3D X-Ray computed tomography data of woven fiber reinforced composites. The main goal of this work is to expand the state of the art in nondestructive testing of composite structures towards lower resolutions and thus larger areas of woven fabrics.

@@

While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}[noitemsep]
\item{ How to extract and visualize the geometric structure of a particular fiber bundle?}
\item{ How to visualize the interaction between a particular pair of fiber bundles (crossovers/braiding) or a unit cell?}
\item{ Which fiber bundles show a particular orientation? }
\item{ Which fiber bundles are of the same type of yarn (i.e. which bundles show similar sizes or diameters), which is the largest or smallest fiber bundle?}
\end{itemize}

Answering the above queries from the volume renderings of XCT datasets, or from visual inspection of particular 2D slices is non-trivial, even for experts. We detail and evaluate a method which uses visualization techniques to gain insight into our data. We demarcate the above queries into two parts: \textit{geometric structure} and \textit{spatial context}. Geometric structure refers to the shape, size, and orientation of a fiber bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non-destructive testing through visualization of composite structures in \textit{lower resolutions} and thus \textit{larger areas} of woven fabrics.
\begin{figure}[tb]
	\centering
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/algo.png}
	%\includegraphics[width=0.5\textwidth,  trim = 0mm 300mm 0mm 0mm, clip,]{imagesMT2014/image_algo}
	\includegraphics[width=0.45\textwidth]{images_pvis/workflow.png}
	\caption{Flow chart of the MetaTracts approach for fiber bundle extraction}
	\label{fig:flowchart}
\end{figure}
Fig.~\ref{fig:flowchart} shows an over view of our approach. Starting from the XCT data, we first generate MetaTracts, cluster the MetaTracts to create fiber bundles and then generate visualizations from the clusters. For a detailed description of the MetaTracts generation process, see Sec.~\ref{subsec:MetaTracts-generation}. For the description of plausible fiber bundle generation see Sec.~\ref{subsec:fiber-bundles}. For a discussion of the visualizations we create for our domain experts see Sec.~\ref{sec:vis}. Sec.~\ref{sec:results} describes our experimental results, Sec.~\ref{sec:param_choices} presents our parameter choices and  Sec.~\ref{sec:user_eval} summarizes our user evaluations.

@@

\subsection {Fiber Clustering in Diffusion Tensor Imaging}
\label {subsec:fiberClus}
In DTI, a similarity measure based on factors such as proximity between fibers is used to group fiber tracts into bundles. Extensive research has been done on automatic DTI fiber clustering methods~ \cite{Brun2004,Brun2003,Corouge2004,westinMEDIA02,Zhang2008}. 
Clustering requires choosing a suitable proximity measure and choosing a method for grouping ``close" fibers.
Pairwise proximity measures include endpoint distances~\cite{Brun2003} and mean of the closest distances between points on two fibers~\cite{Corouge2004}. Zhang et al.~\cite{Zhang2008} introduced a thresholded version of the of mean of closest distances, so that fibers that are close for certain distance and then diverge, are clustered separately. Brun et al.~\cite{Brun2004} uses normalized cuts along with a pairwise distances  measure computed using a 9-D fiber shape descriptors. The choice of the clustering algorithm can be broadly divided into those approaches using hierarchical clustering~\cite{Moberts2005, Zhang2008} and those using spectral clustering~\cite{ Brun2004,jonasson2005, ODonnell2007}.

@@

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results. 
There are a number of difficulties in hierarchical clustering. First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores and deals with fiber extraction from high resolution data where the individual fibers are clearly discernible. Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\includegraphics[width=0.45\textwidth]{images_pvis/data-char.pdf}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Fig. \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Fig.~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Fig.~\ref{fig:data-char}a shows a volume rendering of the dataset. Fig.~\ref{fig:data-char}b shows 2D slices along the $X$- and Fig.~\ref{fig:data-char}c in $Z$-axis. Fig.~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve in the 2D slice images. Fig.~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Fig.~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles. 
We make the following assumptions on our data:
\begin{itemize}[noitemsep]
\item The structure embedded in the data contains fiber bundles of indiscernible fibers.
\item Local orientation: each point within a fiber bundle has a local orientation which is parallel to the corresponding area in the fiber bundle.
\item The local orientation may change gradually inside the fiber bundle.
\item The local orientation may be noisy and not reliable.
\item Connectivity: moving along the direction of a non-noisy local orientation in small increments, we will reach another neighborhood with similar local orientation.
\item Fiber bundles going in different directions only interact near the surface of contact.
\end{itemize}
%\begin{figure}[htp]
%  \centering
%    \subfloat[Overview of the dataset]{\label{figur:3}\includegraphics[scale=0.3]{imagesMetaTracts/crop13-a.eps}}
%  \subfloat[X Axis]{\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/xaxis.eps}}\label{figur:1}\\
%  \subfloat[Y Axis]{\label{figur:2}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/yaxis.eps}}
%  \subfloat[Z Axis]{\label{figur:3}\includegraphics[width=40mm, height=40mm]{imagesMetaTracts/zaxis.eps}}
%  \caption{One of our datasets.Fig \ref{fig:orig_data}(a) provides an overview of the entire volume, the rest show slices of each of the major axis.}\label{fig:orig_data}
%\end{figure}

@@

\section {Extracting MetaTracts }
\label{sec:approach}
Extracting the MetaTracts consists of two  major steps. First we extract a local orientation vector at each grid location (Sec.~\ref{subsec:rh}). Second, we compute coarse poly-cylinders called MetaTracts which traverse the data (Sec.~\ref{subsec:mtprop}).

@@

Grid locations where the $R_{H}$ is above a cutoff threshold are marked as regions with reliable orientations (see Sec.~\ref{sec:param_choices} on parameter choices).

@@

In Fig.~\ref{fig:reliable_hessian} we see the intermediate results of the local orientation computation only at locations where $R_{H}$ is greater than the cutoff threshold. The unit vector representing the principal direction has been mapped to RGB color space. Fig.~\ref{fig:reliable_hessian}a shows the entire data set. Regions where the principal direction is parallel to the X axis are red in color, directions parallel to Y are green and those parallel to the Z are blue. Fig.~\ref{fig:reliable_hessian}b,c show 2D slices along the Z and X axes respectively. Fig.~\ref{fig:reliable_hessian}d shows a magnified region of interest. Note, the dark regions within a bundles are regions where the $R_H$ is less than the threshold and are unreliable. The bundles are not uniformly colored, the Hessians and the corresponding directions are noisy.

@@

%
\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{images_pvis/figure3}
%\includegraphics[width=0.5\textwidth,  trim = 0mm 90mm 0mm 0mm, clip]{images_pvis/figure3}
\includegraphics[width=0.45\textwidth]{images_pvis/reliable_hessian.pdf}
\caption{Reliable Hessians. (a) Colored according to Orientation vector mapped to RGB. (b) 2D slice along Z and (c) along X axis. (d) Magnified region marked in green square (in c).}
\label{fig:reliable_hessian}
\end{figure}

@@

 %glass
We tested our method on three datasets with varying characteristics. We report results on two due to lack of space.  
Dataset 1 is described in Sec.~\ref{sec:char_data}. Dataset 1 has eleven separate fiber bundles which were identified correctly by MetaTracts.
Fig.~\ref{fig:orientation_clustering} shows the result of dataset~\ref{fig:data-char} decomposed into two orientation clusters with each orientation cluster further decomposed into $h=10$ clusters (the ground truth is 6 and 5 respectively).
